---
title: "data cleaning"
author: "Ray"
date: "6/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


Read the data in
```{r}
original <- read_csv('sot_performance_ongoing_waits_dec21.csv',
  col_types = cols(
    MonthEnding = col_character(),
    HBT = col_character(),
    HBTQF = col_character(),
    PatientType = col_character(),
    Specialty = col_character(),
    SpecialtyQF = col_character(),
    NumberWaiting = col_double(),
    NumberWaitingQF = col_character(),
    NumberWaitingOver12Weeks = col_double(),
    NumberWaitingOver12WeeksQF = col_character()
  )
)
```




Select the Inpatient/Day case observations and order it by date
```{r}
original <- original %>%
  filter(PatientType == 'Inpatient/Day case') %>%
  arrange(MonthEnding)

```


Add Year, Month, DaY 
```{r}
a <- original$MonthEnding
Year <- substr(a,1,4) 
Month <- substr(a,5,6) 
Day <- substr(a,7,8)
original <- cbind(original,Year,Month,Day)
```


Create a new column 'Date' to specify the date
```{r}
original <- unite(original, col='Date',remove = FALSE, c('Year', 'Month', 'Day'), sep='-')

original <- original %>%
  mutate(
    Date = as.Date(original$Date)
  )
```




Read the demography data in
```{r}
demography <- read_csv('HBdemography.csv',
  col_types = cols(
    HB = col_character(),
    HBQF = col_character(),
    Sex = col_character(),
    SexQF = col_character()
  )
)


demography.proj <- read_csv('hb_pop_proj_03072020.csv',
  col_types = cols(
    HB = col_character(),
    HBQF = col_character(),
    Sex = col_character(),
    SexQF = col_character()
  )
)

```


Filter the Year to get what we want
```{r}
demography <- demography %>%
  filter(Year >= 2012)

demography.proj <- demography.proj %>%
  filter(Year == 2021)
```

Delete the first column
```{r}
demography <- demography[,-1]
```

Combine the the two demography data set since the demography only have the Year from 2012 to 2020, it doesn't have the Year 2021
```{r}
demography <- rbind(demography,demography.proj)
```


Delete the observations whose HBT code is SB0801
```{r}
original <- original %>%
  filter(HBT != 'SB0801')
```

Change the column name
```{r}
colnames(demography)[2] <- 'HBT'
```


Change the Year and Month variables to be numeric
```{r}
original$Year <- as.numeric(original$Year)
original$Month <- as.numeric(original$Month)
```


all means all the females and males
```{r}
all <- demography %>%
  filter(Sex == 'All')
```


The human age is classified into four categories as Child (0-12 years), Adolescence (13-18 years), Adult (19-59 years) and Senior Adult (60 years and above). Just get the senior.adult
```{r}
Senior.Adult <- all[,67:97]
```

Calculate the Senior.Adult's population
```{r}

Senior.Adult.population <- rowSums(Senior.Adult)
```

calculate the proportion for the Senior.Adult's for every health board in different years
```{r}

Pro.Senior.Adult <- (Senior.Adult.population/all$AllAges)*100
```

Select the variables we want
```{r}
final.demography <- all[,c(1,2,6)]

final.demography <- cbind(final.demography,Pro.Senior.Adult)


```

Merge the two data set by HBT and Year. Then delete the variables we don't want
```{r}
merge.df <- left_join(
  original,
  final.demography,
  by = c('HBT','Year')
)


merge.df <- merge.df[,-c(1,3,4,6,8:10,14)]


```


Order the data set by Date and HBT
```{r}
merge.df <- merge.df %>%
  arrange(Date,HBT)


```

Add to time point variable to the data set 
```{r}
merge.df <- merge.df %>%
  mutate(Time.point = (Year-2012)*12 + Month-10)
```



Create a new variable called Group to specify the HBT and Specialty. And count the observations in every group
```{r}
merge.df <- unite(merge.df, col='Group',remove = FALSE, c('HBT', 'Specialty'), sep='-')



vaild.group <- merge.df %>%
count(Group)


```

Calculate the NA
```{r}
NA.df <- merge.df %>%
  group_by(Group) %>%
  select(NumberWaiting) %>% 
  summarise_all(~sum(is.na(.)))


colnames(NA.df)[2] <- 'num.NA'
```



Join the data
```{r}
NA.merge <- left_join(
  vaild.group,
  NA.df,
  by = c('Group')
)

```


Create a column called vaild.num.n to specify how many available observations per group has
```{r}
vaild.num.n <- NA.merge$n-NA.merge$num.NA

NA.merge <- cbind(NA.merge,vaild.num.n)
```

Delete the groups that don't have any available observations
```{r}
group.have.data <- NA.merge %>%
  filter(vaild.num.n > 0)
```

Find the groups that need to add some rows of zeros
```{r}
group.need.to.add.zeros <- group.have.data %>%
  filter((vaild.num.n + num.NA) < 111)
```

Find the groups that don't need to add some rows of zeros
```{r}
group.not.need.to.add.zeros <- group.have.data %>%
  filter((vaild.num.n + num.NA) == 111)
```

order by the Group variable
```{r}
merge.df <- merge.df %>%
  arrange(Group)
```


filter out the Groups need to add zeros
```{r}
merge.df.trim.group.need.to.add.zeros <- merge.df %>%
  filter(Group %in% group.need.to.add.zeros[,1]) %>%
  arrange(Group)
```

filter out the Groups don't need to add zeros
```{r}
merge.df.trim.group.not.need.to.add.zeros <- merge.df %>%
  filter(Group %in% group.not.need.to.add.zeros[,1]) %>%
  arrange(Group)
```

Create a vector called Group to contain all the Groups' names which are needed to add zeros
```{r}
Group <- group.need.to.add.zeros[,1]
#rep each group 111 times
Group <- rep(Group,each = 111)

```

To make up a data frame that is completed
```{r}
the.make.up.data.frame <- merge.df.trim.group.not.need.to.add.zeros[1:17649,c(5,6,7,10)]

the.make.up.data.frame <- cbind(the.make.up.data.frame,Group)
```


To separate out the Group to HBT and Specialty
```{r}
the.make.up.data.frame <- the.make.up.data.frame %>%
  separate(Group,
           into = c('HBT','Specialty'),
           sep = '-',
           remove = F
           )
  
```

Create a joint.data which is only containing the Date, Group, NumberWaiting.
```{r}
joint.data <- merge.df.trim.group.need.to.add.zeros[,c(5,1,4)]
```

Set the True NA to 1000 first
```{r}
joint.data[is.na(joint.data)] <- 1000
```



To joint the data by Group and Date
```{r}
the.final.make.up.data.frame <- left_join(
  the.make.up.data.frame,
  joint.data,
  by = c('Group','Date')
)
```




Add the zeros on
```{r}
the.final.make.up.data.frame[is.na(the.final.make.up.data.frame)] <- 0
```

Get back NAs
```{r}
the.final.make.up.data.frame$NumberWaiting[the.final.make.up.data.frame$NumberWaiting== 1000] <- NA
```




Join the demography information in
```{r}
the.final.make.up.data.frame <- left_join(
  the.final.make.up.data.frame,
  final.demography,
  by = c('HBT','Year')
)
```




combine the two data frames
```{r}
Final.data <- rbind(merge.df.trim.group.not.need.to.add.zeros,the.final.make.up.data.frame)
```




Find out the groups that have NAs
```{r}
The.groups.have.na <- group.have.data %>%
  filter(num.NA > 0)
```

Choose the groups the have NAs
```{r}
The.data.have.na <- Final.data  %>%
  filter(Group %in% The.groups.have.na[,1])

```

Select the NumberWating out to impute
```{r}
Y <- The.data.have.na$NumberWaiting
```

Library the imputeTS package
```{r}
library(imputeTS)
```

Impute the missing values
```{r}
Y <- round(na_kalman(Y))

The.data.have.na$NumberWaiting <- Y
```


Find out the groups that don't have NAs
```{r}
The.groups.dont.have.na <- group.have.data %>%
  filter(num.NA == 0)
```

Choose the groups the don't have NAs
```{r}
The.data.dont.have.na <- Final.data  %>%
  filter(Group %in% The.groups.dont.have.na[,1])

```


A data frame without NAS
```{r}
Final.data.without.na <- rbind(The.data.dont.have.na,The.data.have.na)
```


preview the data set
```{r}
str(Final.data.without.na)
```




Get the HBT
```{r}
myHBT <- unique(Final.data.without.na$HBT)

myHBT <- myHBT[-14]
```


#EDA


Choose the first 7 HBT but don't contain the Z9 Specialty since it is the aggregated one
```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[1:7]) %>%
  filter(Specialty != 'Z9') %>%
  ggplot(aes(Date, NumberWaiting, color = Specialty)) +
  geom_line() +
  facet_wrap(~ HBT) +
  ggtitle("Figure 3.1 Numer of inpatient/Day case patients waiting in first \n7 HBT for every specialty over time")

```

Choose the last 7 HBT but don't contain the Z9 Specialty since it is the aggregated one
```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[8:14]) %>%
  filter(Specialty != 'Z9') %>%
  ggplot(aes(Date, NumberWaiting, color = Specialty)) +
  geom_line() +
  facet_wrap(~ HBT) +
  ggtitle("Figure 3.2 Numer of inpatient/Day case patients waiting in last \n7 HBT for every specialty over time")
```




```{r}
Final.data.without.na %>%
  filter(Specialty == 'Z9') %>%
  filter(HBT != 'S92000003') %>%
  ggplot(aes(Date, NumberWaiting, color = Group)) +
  geom_line() +
  ggtitle("Figure 3.3 The aggregated numer of inpatient/Day case patients \nwaiting in the 14 HBT over time")
```

```{r}
Final.data.without.na %>% 
  filter(Group == 'S92000003-Z9')  %>%
  ggplot(aes(Date, NumberWaiting, color = Group)) +
  geom_line() +
  ggtitle("Figure 3.4 The aggregated numer of inpatient/Day case patients \nwaiting in the whole Scotland over time")
```


```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[1:7]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(AllAges, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.5 Numer of inpatient/Day case patients waiting in first \n7 HBT V.S population") +
  xlab("Population")
```

```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[8:14]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(AllAges, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.6 Numer of inpatient/Day case patients waiting in last \n7 HBT V.S population") +
  xlab("Population")
```

```{r}
Final.data.without.na %>% 
  filter(Group == 'S92000003-Z9')  %>%
  ggplot(aes(AllAges, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.7 Numer of inpatient/Day case patients waiting in the \nwhole Scotland  V.S population") +
  xlab("Population")
```


```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[1:7]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(Pro.Senior.Adult, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.8 Numer of inpatient/Day case patients waiting in first \n7 HBT V.S the proportion of old people") +
  xlab("Proportion of old people (age >= 60)")
```


```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[8:14]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(Pro.Senior.Adult, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.9 Numer of inpatient/Day case patients waiting in last \n7 HBT V.S the proportion of old people") +
  xlab("Proportion of old people (age >= 60)")
```


```{r}
Final.data.without.na %>% 
  filter(Group == 'S92000003-Z9')  %>%
  ggplot(aes(Pro.Senior.Adult, NumberWaiting, color = Group)) +
  geom_point() +
  ggtitle("Figure 3.10 Numer of inpatient/Day case patients waiting in the \nwhole Scotland  V.S the proportion of old people") +
  xlab("Proportion of old people (age >= 60)")
```

```{r}
Final.data.without.na %>% 
  filter(Group == 'S92000003-Z9')  %>%
  ggplot(aes(x = NumberWaiting)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ .) +
  ggtitle("Figure 3.11 Histogram for the whole Scotland (aggregated level)")
```

```{r}
Final.data.without.na %>% 
  filter(HBT %in% myHBT[1:3]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(x = NumberWaiting)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ ., scales = "free") +
  ggtitle("Figure 3.12 Histogram for the first three health boards (aggregated level)")
```
```{r}
Final.data.without.na %>% 
  filter(HBT %in% myHBT[4:6]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(x = NumberWaiting)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ ., scales = "free") +
  ggtitle("Figure 3.13 Histogram for the second three health boards (aggregated level)")
```


```{r}
Final.data.without.na %>% 
  filter(HBT %in% myHBT[7:9]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(x = NumberWaiting)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ ., scales = "free") +
  ggtitle("Figure 3.14 Histogram for the third three health boards (aggregated level)")
```

```{r}
Final.data.without.na %>% 
  filter(HBT %in% myHBT[10:12]) %>%
  filter(Specialty == 'Z9') %>%
  ggplot(aes(x = NumberWaiting)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ ., scales = "free") +
  ggtitle("Figure 3.15 Histogram for the last three health boards (aggregated level)")
```

#Modeling



```{r}
library("glmmTMB")
```


filter out the Scotland country code and the Z9 aggregated specialty code
```{r}
model.data.without.Z9andS92000003 <- Final.data.without.na %>%
  filter(HBT != 'S92000003') %>%
  filter(Specialty != 'Z9')
  
```

check for overdispersion
```{r}
mean(model.data.without.Z9andS92000003$NumberWaiting)
var(model.data.without.Z9andS92000003$NumberWaiting)
```




create training data set and test data set
```{r}
library(caret)
set.seed(24)
train.idx <- c()


#to make sure every group have 70% of the data as training data 
for(i in 1:318) 
{
group.train.idx <- createDataPartition(model.data.without.Z9andS92000003[(1+(i-1)*111):(i*111),]$NumberWaiting, p = 0.7)$Resample1
group.train.idx <- group.train.idx + (i-1)*111
train.idx <- c(train.idx,group.train.idx)
}



train.data <- model.data.without.Z9andS92000003[train.idx,]
test.data <- model.data.without.Z9andS92000003[-train.idx,]
```



a zero-inflated Poisson model with a single zero inflation parameter applying to all observations
```{r}
fit_zipoisson <- glmmTMB(NumberWaiting~Time.point + Pro.Senior.Adult + offset(log(AllAges)) +(1|HBT/Specialty),
data=train.data,
ziformula=~1,
family=poisson)
```




a standard zero-inflated negative binomial model (variance = µ(1 + µ/k)
```{r}
fit_zinbinom <- update(fit_zipoisson,family=nbinom2)
```


let variance = ϕµ
```{r}
fit_zinbinom1 <- update(fit_zipoisson,family=nbinom1)
```




without offset
```{r}
fit_zinbinom1_bs <- update(fit_zinbinom1,
. ~ Time.point + AllAges + Pro.Senior.Adult  +(1|HBT/Specialty))

```




Hurdle model
```{r}
fit_hnbinom1 <- update(fit_zinbinom1,
ziformula=~.,
data=train.data,
family=truncated_nbinom1)
```



get the prediction for every models for the test data 
```{r}
y_predicted.fit_zipoisson <- predict(fit_zipoisson, test.data, type = "response")
y_predicted.fit_zinbinom <- predict(fit_zinbinom, test.data, type = "response")
y_predicted.fit_zinbinom1 <- predict(fit_zinbinom1, test.data, type = "response")
y_predicted.fit_zinbinom1_bs <- predict(fit_zinbinom1_bs, test.data, type = "response")
y_predicted.fit_hnbinom1 <- predict(fit_hnbinom1, test.data, type = "response")
```



Model comparison
```{r}
data.frame(Model = c("zero-inflated Poisson model","zero-inflated negative binomial model(nbinom2)","zero-inflated negative binomial model(nbinom1)","zero-inflated negative binomial model(nbinom1) without offset","Hurdle negative binomial model(nbinom1)"),
  RMSE = c(RMSE(y_predicted.fit_zipoisson, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1, test.data$NumberWaiting)),
            MAE = c(MAE(y_predicted.fit_zipoisson, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1, test.data$NumberWaiting)),
  AIC =c(AIC(fit_zipoisson),AIC(fit_zinbinom),AIC(fit_zinbinom1),AIC(fit_zinbinom1_bs),AIC(fit_hnbinom1)))
```







The summary output
```{r}
summary(fit_hnbinom1)
```





without HBT random effect
```{r}
fit_hnbinom1.without.HBT.random.effect <- update(fit_hnbinom1,
                                                 . ~ Time.point + Pro.Senior.Adult + offset(log(AllAges)) +(1|HBT:Specialty),
ziformula=~.,
data=train.data,
family=truncated_nbinom1)
```



To test if the HBT random effect important
```{r}
anova(fit_hnbinom1.without.HBT.random.effect,fit_hnbinom1)
```

```{r}
summary(fit_hnbinom1.without.HBT.random.effect)
```





Without pro.senior.adult
```{r}
fit_hnbinom1.without.pro.senior.adult <- update(fit_hnbinom1.without.HBT.random.effect,
                                                 . ~ Time.point  + offset(log(AllAges)) +(1|HBT:Specialty))
```

AIC value
```{r}
AIC(fit_hnbinom1.without.pro.senior.adult)
```

test if Pro.Senior.Adult is significant
```{r}
anova(fit_hnbinom1.without.HBT.random.effect,fit_hnbinom1.without.pro.senior.adult)
```

prediction for the new two models
```{r}
y_predicted.fit_hnbinom1.without.HBT.random.effect <- predict(fit_hnbinom1.without.HBT.random.effect, test.data, type = "response")
y_predicted.fit_hnbinom1.without.pro.senior.adult <- predict(fit_hnbinom1.without.pro.senior.adult, test.data, type = "response")
```


model comparison
```{r}
data.frame(Model = c("zero-inflated Poisson model","zero-inflated negative binomial model(nbinom2)","zero-inflated negative binomial model(nbinom1)","zero-inflated negative binomial model(nbinom1) without offset","Hurdle negative binomial model(nbinom1)","no HBT random effect", "no HBT random effect and no Pro.Senior.Adult"),
  RMSE = c(RMSE(y_predicted.fit_zipoisson, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1.without.HBT.random.effect, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1.without.pro.senior.adult, test.data$NumberWaiting)),
            MAE = c(MAE(y_predicted.fit_zipoisson, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1.without.HBT.random.effect, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1.without.pro.senior.adult, test.data$NumberWaiting)),
  AIC =c(AIC(fit_zipoisson),AIC(fit_zinbinom),AIC(fit_zinbinom1),AIC(fit_zinbinom1_bs),AIC(fit_hnbinom1),AIC(fit_hnbinom1.without.HBT.random.effect),AIC(fit_hnbinom1.without.pro.senior.adult)))
```



fit a model without poisson
```{r}
fit.without.link.family <- glmmTMB(NumberWaiting~Time.point + Pro.Senior.Adult + AllAges +(1|HBT:Specialty),
data=train.data)
```


prediction for the model without link family
```{r}
y_predicted.fit.without.link.family <- predict(fit.without.link.family, test.data, type = "response")
```


model comparison
```{r}
data.frame(Model = c("zero-inflated Poisson model","zero-inflated negative binomial model(nbinom2)","zero-inflated negative binomial model(nbinom1)","zero-inflated negative binomial model(nbinom1) without offset","Hurdle negative binomial model(nbinom1)","no HBT random effect", "no Pro.Senior.Adult","model without any link family(lmm model)"),
  RMSE = c(RMSE(y_predicted.fit_zipoisson, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),RMSE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1.without.HBT.random.effect, test.data$NumberWaiting),RMSE(y_predicted.fit_hnbinom1.without.pro.senior.adult, test.data$NumberWaiting),RMSE(y_predicted.fit.without.link.family,test.data$NumberWaiting)),
            MAE = c(MAE(y_predicted.fit_zipoisson, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1, test.data$NumberWaiting),MAE(y_predicted.fit_zinbinom1_bs, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1.without.HBT.random.effect, test.data$NumberWaiting),MAE(y_predicted.fit_hnbinom1.without.pro.senior.adult, test.data$NumberWaiting),MAE(y_predicted.fit.without.link.family,test.data$NumberWaiting)),
  AIC =c(AIC(fit_zipoisson),AIC(fit_zinbinom),AIC(fit_zinbinom1),AIC(fit_zinbinom1_bs),AIC(fit_hnbinom1),AIC(fit_hnbinom1.without.HBT.random.effect),AIC(fit_hnbinom1.without.pro.senior.adult),AIC(fit.without.link.family)))
```






The diagnostic is ok
```{r}
qqnorm(ranef(fit_hnbinom1.without.HBT.random.effect)$cond$"HBT:Specialty"[[1]],main="HBT:Specialty effects")
```



Observed values v.s predicited values
```{r}

plot(test.data$NumberWaiting,main='Observed v.s predicted',xlab='Index', ylab='Response',col='green')
par(new=TRUE)
plot(y_predicted.fit_hnbinom1.without.HBT.random.effect,axes= FALSE, xlab='', ylab='',col='red')
legend("topright",c("Observed values","Predicted values"),cex=1,col=c("green","red"),pch=c(1,1))
```






create a data frame for Observed values in test data and predicted values of the best model
```{r}
data_mod <- data.frame(Predicted = y_predicted.fit_hnbinom1.without.HBT.random.effect,
                       Observed = test.data$NumberWaiting
                       )
```



Observed values in test data V.S predicted values by the best model
```{r}
ggplot(data_mod,
       aes(x = Predicted,
           y = Observed)) +
  geom_point() +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              size = 2) + 
    ggtitle("Observed values in test data V.S predicted values by the best model")
  
```



summary output for the best model
```{r}
summary(fit_hnbinom1.without.HBT.random.effect)

```


The random effects' absolute intercepts below 0.2 for Conditional model
```{r}
a <- ranef(fit_hnbinom1.without.HBT.random.effect, condVar=TRUE)
a <- a[[1]]
a <- data.frame(a)

a<- a %>%
  filter(X.Intercept. < 0.2) %>%
  filter(X.Intercept. > -0.2)



dotchart(a$X.Intercept., labels = rownames(a), pch = 21, bg = "green", pt.cex = 1.5,main="The random effects' absolute intercepts below 0.2 \nfor Conditional model ")



```

The random effects' absolute intercepts below 0.2 for Zero-inflation model
```{r}
b <- ranef(fit_hnbinom1.without.HBT.random.effect, condVar=TRUE)
b <- b[[2]]
b <- data.frame(b)

b<- b %>%
  filter(X.Intercept. < 0.2) %>%
  filter(X.Intercept. > -0.2)



dotchart(b$X.Intercept., labels = rownames(a), pch = 21, bg = "green", pt.cex = 1.5,main="The random effects' absolute intercepts below 0.2 \nfor Zero-inflation model ")
```



#Specific-to-specific model



Just look at the group S080000032-C8
```{r}
Final.data.without.na %>%
  filter(HBT %in% myHBT[13]) %>%
  filter(Specialty == c('C8') )%>%
  ggplot(aes(Date, NumberWaiting, color = Group)) +
  geom_line() +
  ggtitle("Numer of inpatient/Day case patients waiting in the group \nS080000032-C8 over time")
```

select this group's training data and test data
```{r}
one.group.data <- Final.data.without.na %>%
  filter(HBT %in% myHBT[13]) %>%
  filter(Specialty == c('C8')) 

one.group.train.data <- train.data %>% 
  filter(HBT %in% myHBT[13]) %>%
  filter(Specialty == c('C8'))

one.group.test.data <- test.data %>% 
  filter(HBT %in% myHBT[13]) %>%
  filter(Specialty == c('C8'))

```

check for dispersion
```{r}
mean(one.group.data$NumberWaiting)
var(one.group.data$NumberWaiting)
```


```{r}
fit.qpois <- glm(NumberWaiting~Time.point + Pro.Senior.Adult + offset(log(AllAges)), data=one.group.train.data, family=quasipoisson)
summary(fit.qpois)

library(MASS)
fit.negbin1 <-glm.nb(NumberWaiting~Time.point + Pro.Senior.Adult + offset(log(AllAges)), data=one.group.train.data)
summary(fit.negbin1)
```


check for significance of offset
```{r}
fit.negbin1.without.offset <-glm.nb(NumberWaiting~Time.point + Pro.Senior.Adult + AllAges, data=one.group.train.data)
summary(fit.negbin1.without.offset)
```

check for significance of offset
```{r}
fit.qpois.without.offset <- glm(NumberWaiting~Time.point + Pro.Senior.Adult + AllAges, data=one.group.train.data, family=quasipoisson)
summary(fit.qpois.without.offset)
```
check for significance of pro.senior.adult
```{r}
fit.qpois.without.offset.pro.senior.adult <- glm(NumberWaiting~Time.point + AllAges, data=one.group.train.data, family=quasipoisson)
summary(fit.qpois.without.offset.pro.senior.adult)
```
check for significance of pro.senior.adult
```{r}
fit.negbin1.without.offset.pro.senior.adult <-glm.nb(NumberWaiting~Time.point + AllAges, data=one.group.train.data)
summary(fit.negbin1.without.offset.pro.senior.adult)
```

AIC for all the potential models
```{r}
AIC(fit.qpois)
AIC(fit.qpois.without.offset)
AIC(fit.qpois.without.offset.pro.senior.adult)
AIC(fit.negbin1)
AIC(fit.negbin1.without.offset)
AIC(fit.negbin1.without.offset.pro.senior.adult)
```

predcited value for test data
```{r}
fit.qpois.pred <- predict(fit.qpois, one.group.test.data, type = "response")
fit.qpois.without.offset.pred <- predict(fit.qpois.without.offset, one.group.test.data, type = "response")
fit.qpois.without.offset.pro.senior.adult.pred <- predict(fit.qpois.without.offset.pro.senior.adult, one.group.test.data, type = "response")
fit.negbin1.pred <- predict(fit.negbin1, one.group.test.data, type = "response")
fit.negbin1.without.offset.pred <- predict(fit.negbin1.without.offset, one.group.test.data, type = "response")
fit.negbin1.without.offset.pro.senior.adult.pred <- predict(fit.negbin1.without.offset.pro.senior.adult, one.group.test.data, type = "response")

```

RMSES all the potential models
```{r}
RMSE(fit.qpois.pred,one.group.test.data$NumberWaiting)
RMSE(fit.qpois.without.offset.pred,one.group.test.data$NumberWaiting)
RMSE(fit.qpois.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)
RMSE(fit.negbin1.pred,one.group.test.data$NumberWaiting)
RMSE(fit.negbin1.without.offset.pred,one.group.test.data$NumberWaiting)
RMSE(fit.negbin1.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)
```

MAES all the potential models
```{r}
MAE(fit.qpois.pred,one.group.test.data$NumberWaiting)
MAE(fit.qpois.without.offset.pred,one.group.test.data$NumberWaiting)
MAE(fit.qpois.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)
MAE(fit.negbin1.pred,one.group.test.data$NumberWaiting)
MAE(fit.negbin1.without.offset.pred,one.group.test.data$NumberWaiting)
MAE(fit.negbin1.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)
```


diagnostic for the best two models
```{r}
library(statmod)
par(mfrow=c(1,2))
qqnorm(qresid(fit.qpois.without.offset.pro.senior.adult)); qqline(qresid(fit.qpois.without.offset.pro.senior.adult))
qqnorm(qresid(fit.negbin1.without.offset.pro.senior.adult)); qqline(qresid(fit.negbin1.without.offset.pro.senior.adult))
```
prediction by the glmm model
```{r}
fit_hnbinom1.without.HBT.random.effect.pred <- predict(fit_hnbinom1.without.HBT.random.effect, one.group.test.data, type = "response")
```

model comparison
```{r}
data.frame(Model = c("best glmm model","Quasi-Poisson model","negative binomial model"),
  RMSE = c(RMSE(fit_hnbinom1.without.HBT.random.effect.pred, one.group.test.data$NumberWaiting),RMSE(fit.qpois.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting),RMSE(fit.negbin1.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)),
            MAE = c(MAE(fit_hnbinom1.without.HBT.random.effect.pred, one.group.test.data$NumberWaiting),MAE(fit.qpois.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting),MAE(fit.negbin1.without.offset.pro.senior.adult.pred,one.group.test.data$NumberWaiting)),
  AIC =c(AIC(fit_hnbinom1.without.HBT.random.effect),AIC(fit.qpois.without.offset.pro.senior.adult),AIC(fit.negbin1.without.offset.pro.senior.adult)))
```


Observed values v.s predicted values
```{r}
par(mfrow=c(1,3))
plot(one.group.test.data$NumberWaiting,main='Observed v.s predicted \nof the best glmm model',xlab='Index', ylab='Response',col='green')
par(new=TRUE)
plot(fit_hnbinom1.without.HBT.random.effect.pred,axes= FALSE, xlab='', ylab='',col='red')
legend("topleft",c("Observed values","Predicted values"),cex=1,col=c("green","red"),pch=c(1,1))

plot(one.group.test.data$NumberWaiting,main='Observed v.s predicted \nof the Quasi-Poisson model',xlab='Index', ylab='Response',col='green')
par(new=TRUE)
plot(fit.qpois.without.offset.pro.senior.adult.pred,axes= FALSE, xlab='', ylab='',col='red')
legend("topleft",c("Observed values","Predicted values"),cex=1,col=c("green","red"),pch=c(1,1))

plot(one.group.test.data$NumberWaiting,main='Observed v.s predicted \nof the negative binomial model',xlab='Index', ylab='Response',col='green')
par(new=TRUE)
plot(fit.negbin1.without.offset.pro.senior.adult.pred,axes= FALSE, xlab='', ylab='',col='red')
legend("topleft",c("Observed values","Predicted values"),cex=1,col=c("green","red"),pch=c(1,1))
```






# Need to be contiued, job is not finsihed ...



nest the data by HBT and Specialty
```{r}
by_hbt_specialty <- Final.data.without.na %>% 
  group_by(HBT, Specialty) %>% 
  nest()


```







